Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-19 paper/v2/paper-edited.tex
Copy Edits
Corrections made to maintain grammatical word usage,  sentence
structure and clarity of meaning.

\textbf{Abstract---}
We develop \sysname, a language and compiler to help network operators with a challenging, error-prone task---bridging the gap between network-wide routing objectives and low-level configurations of devices that run complex, distributed protocols.
%
The language allows operators to specify their objectives naturally, using high-level constraints and ordered preferences on paths that traffic should take.
%through their network or other networks.
%using ranked sets of paths which may be partial or end-to-end and may also refer to other networks.
%
The compiler automatically translates these specifications to router-level BGP configurations, using a novel data structure that compactly encodes the flow of routing information along policy-compliant paths.
%It uses We introduce data structures and algorithms that
It guarantees that the compiled configurations correctly implement the specified policy under all possible combinations of failures.
%
We show that \sysname can effectively express the policies of datacenter and backbone networks of a large cloud provider;  and despite its strong guarantees, our compiler scales to networks with hundreds or thousands of routers.

@@

However, this implementation has incorrect behavior in the face of failures. Suppose links X--G and X--H fail. Then, X will hear announcements for \CD{PL*} from C and D, having traversed from G and H to Y to C and D. Per its policy implementation, X will start "leaking" these prefixes externally. Depending on the rationale for local services, this leak could impact security (e.g., if the services are sensitive) or availability (e.g., if the \CD{PL*} prefixes are reused for other services outside of the datacenter). This problem does not manifest without failures because then X has and prefers paths to \CD{PL*} through G and H since they are shorter. A similar problem will happen if links Y--G and Y--H fail.
%\footnote{A different problem occurs when links X--C and X--D (or Y--C and Y--D) fail. X (Y) may stop announcing the global prefixes because they would be heard through G and H.}
Link failures in datacenters are frequent and it is not uncommon to have multiple failed links at a given time~\cite{dc-failure-study}.

@@

Since, by default, routing allows transit traffic (\EG, traffic entering from
\CD{Peer} and leaving through \CD{Prov}), we separately define a
policy to enforce P2 and P5 from Figure~\ref{fig:example1}, using conjunction (\AND),
disjunction (\OR) and negation (\NOT) of constraints.
First, we create reusable abstractions for describing traffic that transits our network. In \sysname, this is done by creating a new parameterized definition.
\begin{code}
\Define transit(X,Y) = \Enter(X{\OR}Y) \AND \Exit(X{\OR}Y)
\Define cust-transit(X,Y) = \Later(X) \AND \Later(Y)
\end{code}
\noindent
Here we define transit traffic between groups of neighbors $X$ and $Y$ as traffic that enters the network through some neighbor in $X$ or $Y$ and then leaves the network through some neighbor in either $X$ or $Y$. Similarly, we define customer transit for customer $X$ and provider $Y$ as traffic that later goes through both $X$ and $Y$ after leaving our network. Using these two new abstractions, we can now implement policies P2 and P5 with the following constraint.

@@

%% \begin{lstlisting}[mathescape=true]
%% define Routing = {
%%     $\path{p_{g*}}{any}$
%%     $\path{p_{l*}}{\NOT enter(out)}$
%%     $\path{true}{exit(out)}$
%% }
%% \end{lstlisting}
\begin{code}
\Define Locality =
    \{PL1 \OR PL2 \Path \textbf{only}(\In)\}
\end{code}
\noindent
%The first line states that there is no restriction (\Any)
%on how traffic must traverse the network for global prefixes.
%aside from the default restriction
%that traffic must not pass through the user's network and then loop
%back on itself.
%This means traffic for
%\CD{PG*} may be sent either from other datacenter routers or
%from external ASes.
This definition says that traffic for local
prefixes only follows paths that are internal to the network at each hop.
This constraint guarantees that the services remain reachable only to locations
inside the datacenter.

@@

To handle these challenges, we decompose compilation into multiple stages, shown in  Figure~\ref{fig:pipeline}, and develop efficient algorithms for the translation between stages. The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for well-formedness (\EG, never constraining traffic that does not enter the network), before being combined with the network topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines safety, it translates the PGIR to an abstract BGP (ABGP) representation. ABGP can then be translated into various vendor-specific device
configurations as needed.
%The \sysname compiler currently generates
% Quagga router configurations from ABGP.

@@

Intuitively, PGIR tracks which states of each automaton the policy is in as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$, and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $v$ such that the preference function $P(v) = \set{i_1, \dots, i_m}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_m$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possiblity is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton can not accept the current path or any extension of it. Since node $(W,5,-)$ is in an accepting state for the first automata, it indicates that this path has preference 1.

\para{Minimization}
After building the PGIR as desribed above,
we minimize it in order to improve the subsequent analysis that checks if the policies captured by it are safe under failures.
The minimization is based on the observation that, although every path in the PGIR is a valid path in the topology, we do not want to consider paths that form loops. In particular, BGP's loop prevention mechanism forces an AS to reject any route for which it is already in the AS path.
%
For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
The precision of the failure-safety analysis often increases when nodes and edges that are not possible due to loops are eliminated.
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.

@@

A natural question then is how the compiler should order PGIR contexts for each router in a way that is both \emph{sound} and \emph{complete} under all possible failures? To this end, we introduce the notion of \emph{regret-free} preferences, motivated by the observations from the example in Figure~\ref{fig:unimplementable}. Router $r$ has a \emph{regret-free} preference for a set of advertisements $X$ over $Y$ if, whevever $r$ selects an advertisement from $X$ over another from $Y$, there is always some policy-compliant path to each final (advertisement) destination $d$ that is at least as good $(\leq_{rank})$ as any possible path to $d$ if it had accepted an advertisement from $Y$ instead. In other words, the preference of $X$ over $Y$ at $r$ is regret-free if we are never worse off by choosing from $X$.
%
In the compilation example from Figure~\ref{fig:example-compilation}, the choice for $C$ to prefer context $(C,3,2)$ to $(C,-,2)$ is regret-free, since there will always be at least as good a path to destination $W$ regardless of any failures that might occur in the network.

@@

After configuration generation, the compiler further processes the ABGP policy, removing community tags when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neigbor the route is being imported from. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

@@

We obtained routing policy for the backbone network and for the datacenters of a large cloud provider. Multiple datacenters share this policy. The backbone network connects to the datacenters and also has many external BGP neighbors. The high-level policies of these networks are captured in an English document which guides operators when writing configuration templates for datacenter routers or actual configurations for the backbone network (where templates are not used because the network has less regular structure).

@@

We found that we could translate all network policies to \sysname. We verified with the operators that our translation preserved intended semantics.\footnote{Not intended as a scientific test, but we also asked the two operators if they would find it easy to express their policies in \sysname. The datacenter operator said that he found the language intuitive. The backbone operator said that formalizing the policy in \sysname seemed equally easy or difficult as formalizing in RPSL~\cite{RFC2622}, but he appreciated that he would have to do it only once for the whole network (not per-router) and did not have to manually compute various local preferences, import-export filters, and MEDs.} We found that the datacenter policies were correctly translated. For the backbone network, the operator mentioned an additional policy that was not present in the English document, which we added later.

>>>>

\textbf{Abstract---}
We develop \sysname, a language and compiler to help network operators with a challenging, error-prone task---bridging the gap between network-wide routing objectives and low-level configurations of devices that run complex, distributed protocols.
%
The language allows operators to specify their objectives naturally, using high-level constraints on and ordered preferences for paths that traffic should take.
%through their network or other networks.
%using ranked sets of paths which may be partial or end-to-end and may also refer to other networks.
%
The compiler automatically translates these specifications to router-level BGP configurations, using a novel data structure that compactly encodes the flow of routing information along policy-compliant paths.
%It uses We introduce data structures and algorithms that
It guarantees that the compiled configurations correctly implement the specified policy under all possible combinations of failures.
%
We show that \sysname can effectively express the policies of datacenter and backbone networks of a large cloud provider;  and despite its strong guarantees, our compiler scales to networks with hundreds or thousands of routers.

@@

However, this implementation has incorrect behavior in the face of failures. Suppose links X--G and X--H fail. Then, X will hear announcements for \CD{PL*} from C and D, having traversed from G and H to Y to C and D. Per its policy implementation, X will start "leaking" these prefixes externally. Depending on the rationale for local services, this leak could impact security (e.g., if the services are sensitive) or availability (e.g., if the \CD{PL*} prefixes are reused for other services outside of the datacenter). This problem does not manifest without failures because then X has and prefers paths to \CD{PL*} through G and H since they are shorter. A similar problem will occur if links Y--G and Y--H fail.
%\footnote{A different problem occurs when links X--C and X--D (or Y--C and Y--D) fail. X (Y) may stop announcing the global prefixes because they would be heard through G and H.}
Link failures in datacenters are frequent and it is not uncommon to have multiple failed links at a given time~\cite{dc-failure-study}.

@@

Since, routing allows transit traffic by default (\EG, traffic entering from
\CD{Peer} and leaving through \CD{Prov}), we separately define a
policy to enforce P2 and P5 from Figure~\ref{fig:example1}, using conjunction (\AND),
disjunction (\OR) and negation (\NOT) of constraints.
First, we create reusable abstractions for describing traffic that transits our network. In \sysname, this is done by creating a new parameterized definition.
\begin{code}
\Define transit(X,Y) = \Enter(X{\OR}Y) \AND \Exit(X{\OR}Y)
\Define cust-transit(X,Y) = \Later(X) \AND \Later(Y)
\end{code}
\noindent
Here we define transit traffic between groups of neighbors $X$ and $Y$ as traffic that enters the network through some neighbor in $X$ or $Y$ and then leaves the network through some neighbor in either $X$ or $Y$. Similarly, we define customer transit for customer $X$ and provider $Y$ as traffic that later goes through both $X$ and $Y$ after leaving our network. Using these two new abstractions, we can now implement policies P2 and P5 with the following constraint.

@@

%% \begin{lstlisting}[mathescape=true]
%% define Routing = {
%%     $\path{p_{g*}}{any}$
%%     $\path{p_{l*}}{\NOT enter(out)}$
%%     $\path{true}{exit(out)}$
%% }
%% \end{lstlisting}
\begin{code}
\Define Locality =
    \{PL1 \OR PL2 \Path \textbf{only}(\In)\}
\end{code}
\noindent
%The first line states that there is no restriction (\Any)
%on how traffic must traverse the network for global prefixes.
%aside from the default restriction
%that traffic must not pass through the user's network and then loop
%back on itself.
%This means traffic for
%\CD{PG*} may be sent either from other datacenter routers or
%from external ASes.
This definition says that traffic for local
prefixes only follows paths that are internal to the network at each hop.
This constraint guarantees that the services remain accessible only to locations
inside the datacenter.

@@

To handle these challenges, we decompose compilation into multiple stages, shown in  Figure~\ref{fig:pipeline}, and develop efficient algorithms for the translation between stages. The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for being well-formed (\EG, never constraining traffic that does not enter the network), before being combined with the network topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines safety, it translates the PGIR to an abstract BGP (ABGP) representation. ABGP can then be translated into various vendor-specific device
configurations as needed.
%The \sysname compiler currently generates
% Quagga router configurations from ABGP.

@@

Intuitively, PGIR tracks the policy states of each automaton as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$, and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $v$ such that the preference function $P(v) = \set{i_1, \dots, i_m}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_m$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possiblity is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton can not accept the current path or any extension of it. Since node $(W,5,-)$ is in an accepting state for the first automata, it indicates that this path has preference 1.

\para{Minimization}
After building the PGIR as desribed above,
we minimize it in order to improve the subsequent analysis that checks if the policies captured by it are safe under failures.
The minimization is based on the observation that, although every path in the PGIR is a valid path in the topology, we do not want to consider paths that form loops. In particular, BGP's loop prevention mechanism forces an AS to reject any route that is already in the AS path.
%
For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
The precision of the failure-safety analysis often increases when nodes and edges that are not possible due to loops are eliminated.
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.

@@

A natural question then is, how should the compiler order PGIR contexts for each router in a way that is both \emph{sound} and \emph{complete} under all possible failures? To this end, we introduce the notion of \emph{regret-free} preferences, motivated by the observations from the example in Figure~\ref{fig:unimplementable}. Router $r$ has a \emph{regret-free} preference for a set of advertisements $X$ over $Y$ if, whevever $r$ selects an advertisement from $X$ over another from $Y$, there is always some policy-compliant path to each final (advertisement) destination $d$ that is at least as good $(\leq_{rank})$ as any possible path to $d$ if it had accepted an advertisement from $Y$ instead. In other words, the preference of $X$ over $Y$ at $r$ is regret-free if we are never worse off by choosing from $X$.
%
In the compilation example from Figure~\ref{fig:example-compilation}, the choice for $C$ to prefer context $(C,3,2)$ to $(C,-,2)$ is regret-free, since there will always be at least as good a path to destination $W$ regardless of any failures that might occur in the network.

@@

After configuration generation, the compiler further processes the ABGP policy, removing community tags when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neighbor from which the route is being imported. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

@@

We obtained routing policy for the backbone network and for the datacenters of a large cloud provider. Multiple datacenters share this policy. The backbone network connects to the datacenters and also has many external BGP neighbors. The high-level policies of these networks are captured in an English document which guides operators when writing configuration templates for datacenter routers or actual configurations for the backbone network (where templates are not used because the network has a less regular structure).

@@

We found that we could translate all network policies to \sysname. We verified with the operators that our translation preserved intended semantics.\footnote{Not intended as a scientific test, but we also asked the two operators if they would find it easy to express their policies in \sysname. The datacenter operator said that he found the language intuitive. The backbone operator said that formalizing the policy in \sysname seemed equally easy or difficult as formalizing in RPSL~\cite{RFC2622}, but he appreciated that he would have to do it only once for the whole network (not per-router) and did not have to manually compute various local preferences, import-export filters, and MEDs.} We further found that the datacenter policies were correctly translated. For the backbone network, the operator mentioned an additional policy that was not present in the English document, which we added later.
